Supplementary Material for Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-04-11 supp_mat_complexity_selection.Rmd
exploring effects of plant genotype on selection gradients.

Load dataset. This dataset is a subset that only includes larva that survive to pupation or turned into parasitoids. We omitted any data where larva mortality was unclear. This enabled us to focus on the selection pressures imposed by insect parasitoids rather than other sources.
```{r Preliminaries, message=FALSE, warning=FALSE, include=FALSE}
## LOAD REQUIRED LIBRARIES & SET OPTIONS ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gsg) # calculating selection gradients

@@

```{r}
size_models <- gall_selection.df %>%
  group_by(Treatment.focus, Genotype) %>%
  do(model = lm(gall_survival ~ Gall_Height_mm, data = .)) %>%
  tidy(model) %>%
  ungroup()

size_models %>%
  ggplot(., aes(x=Genotype, y=estimate, fill=Treatment.focus)) +
  geom_bar(stat="identity", position = "dodge") +
  geom_errorbar(aes(ymax = estimate + std.error, ymin = estimate - std.error), #, color=Treatment.focus
                width = 0.5, position = position_dodge(width = 0.8)) +
  facet_wrap(~term, scales = "free_y")
```

```{r}
size.grad_models <- gall_selection.df %>%
  group_by(Treatment.focus, Genotype) %>%
  do(model = lm(gall_survival/mean(gall_survival) ~ I(Gall_Height_mm - mean(Gall_Height_mm)), data = .)) %>%
  tidy(model) %>%
  ungroup()
size.grad_models

size.grad_models %>%
  filter(term != "(Intercept)") %>%
  ggplot(., aes(x=Genotype, y=estimate, fill=Treatment.focus)) +
  geom_bar(stat="identity", position = "dodge") +
  geom_errorbar(aes(ymax = estimate + std.error, ymin = estimate - std.error), #, color=Treatment.focus
                width = 0.5, position = position_dodge(width = 0.8))

size.grad_models %>%
  filter(term != "(Intercept)") %>%
  ggplot(., aes(x=Treatment.focus, y=estimate, fill=Genotype, group=Genotype)) +
  geom_line() +
  geom_point(shape=21, position = "dodge", size=4) +
  geom_hline(yintercept=0, linetype="dotted")

summary(lm(estimate ~ Treatment.focus, data = filter(size.grad_models, term != "(Intercept)")))
```

>>>>

Load dataset. This dataset is a subset that only includes larva that survive to pupation or turned into parasitoids. We omitted any data where larva mortality was unclear. This enabled us to focus on the selection pressures imposed by insect parasitoids rather than other sources.
```{r Preliminaries, message=FALSE, warning=FALSE, include=FALSE}
## LOAD REQUIRED LIBRARIES & SET OPTIONS ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gsg) # calculating selection gradients
library(visreg)

@@

Analyze separate models for absolute fitness as a function of gall size for each treatment and genotype combination. 
```{r}
size_models <- gall_selection.df %>%
  group_by(Treatment.focus, Genotype) %>%
  do(model = lm(gall_survival ~ I(Gall_Height_mm-mean(Gall_Height_mm)), data = .)) %>%
  tidy(model) %>%
  ungroup()

size_models %>%
  ggplot(., aes(x=Genotype, y=estimate, fill=Treatment.focus)) +
  geom_errorbar(aes(ymax = estimate + std.error, ymin = estimate - std.error), #, color=Treatment.focus
                width = 0.5, position = position_dodge(width=0.8)) +
  geom_point(shape=21, size = 2, position = position_dodge(width=0.8)) +
  geom_hline(yintercept = 0, linetype="dotted") +
  facet_wrap(~term, scales = "free_y")
```

```{r}
size.geno <- gall_selection.df %>%
  group_by(Treatment.focus, Genotype) %>%
  summarise_at(vars(Gall_Height_mm), funs(mean))
size.geno

size.grad_models <- gall_selection.df %>%
  group_by(Treatment.focus, Genotype) %>%
  do(model = lm(gall_survival/mean(gall_survival) ~ I(Gall_Height_mm - mean(Gall_Height_mm)), data = .)) %>%
  tidy(model) %>%
  ungroup()
size.grad_models

size.grad_models %>%
  filter(term != "(Intercept)") %>%
  ggplot(., aes(x=Genotype, y=estimate, fill=Treatment.focus)) +
  geom_errorbar(aes(ymax = estimate + std.error, ymin = estimate - std.error), #, color=Treatment.focus
                width = 0.5, position = position_dodge(width = 0.8)) +
  geom_point(shape=21, size = 2, position = position_dodge(width=0.8)) +
  geom_hline(yintercept = 0, linetype="dotted")

size.grad_models %>%
  filter(term != "(Intercept)") %>%
  ggplot(., aes(x=Treatment.focus, y=estimate, fill=Genotype, group=Genotype)) +
  geom_line() +
  geom_point(shape=21, size=4) +
  geom_hline(yintercept=0, linetype="dotted")

gall_selection.df$c.Gall_Height_mm <- with(gall_selection.df, Gall_Height_mm - mean(Gall_Height_mm))

size_GxE_lms <- lm(gall_survival ~ (c.Gall_Height_mm+Genotype+Treatment.focus)^2, data = gall_selection.df, contrast = list(Genotype = "contr.sum"))
summary(size_GxE_lms)
anova(size_GxE_lms)

table(gall_selection.df$Genotype, gall_selection.df$Treatment.focus)

test <- size.grad_models %>%
  filter(term != "(Intercept)") %>%
  left_join(., size.geno)

cor.test(test$estimate, test$Gall_Height_mm)
```

```{r}
gall_agg <- gall_selection.df %>%
  group_by(Genotype, Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise_at(vars(gall_individuals, gall_survival), funs(mean)) %>%
  ungroup() %>%
  mutate(c.gall_individuals = gall_individuals - mean(gall_individuals))

ind_models <- gall_agg %>%
  group_by(Treatment.focus, Genotype) %>%
  do(model = lm(gall_survival ~ I(gall_individuals-mean(gall_individuals)), data = .)) %>%
  tidy(model) %>%
  ungroup()

ind_models %>%
  ggplot(., aes(x=Genotype, y=estimate, fill=Treatment.focus)) +
  geom_errorbar(aes(ymax = estimate + std.error, ymin = estimate - std.error), #, color=Treatment.focus
                width = 0.5, position = position_dodge(width=0.8)) +
  geom_point(shape=21, size = 2, position = position_dodge(width=0.8)) +
  geom_hline(yintercept = 0, linetype="dotted") +
  facet_wrap(~term, scales = "free_y")

ind.grad_models <- gall_agg %>%
  group_by(Treatment.focus, Genotype) %>%
  do(model = lm(gall_survival/mean(gall_survival) ~ I(gall_individuals - mean(gall_individuals)), data = .)) %>%
  tidy(model) %>%
  ungroup()
ind.grad_models

ind.grad_models %>%
  filter(term != "(Intercept)") %>%
  ggplot(., aes(x=Genotype, y=estimate, fill=Treatment.focus)) +
  geom_errorbar(aes(ymax = estimate + std.error, ymin = estimate - std.error), #, color=Treatment.focus
                width = 0.5, position = position_dodge(width = 0.8)) +
  geom_point(shape=21, size = 2, position = position_dodge(width=0.8)) +
  geom_hline(yintercept = 0, linetype="dotted") 
```

```{r}
plant_agg <- gall_selection.df %>%
  group_by(Genotype, Treatment.focus, Plant_Position) %>%
  summarise_at(vars(Density_per_100_shoots, gall_survival), funs(mean)) %>%
  ungroup() %>%
  mutate(c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

density.grad_models <- plant_agg %>%
  group_by(Treatment.focus, Genotype) %>%
  do(model = lm(gall_survival/mean(gall_survival) ~ I(Density_per_100_shoots - mean(Density_per_100_shoots)), data = .)) %>%
  tidy(model) %>%
  ungroup()
density.grad_models

density.grad_models %>%
  filter(term != "(Intercept)") %>%
  ggplot(., aes(x=Genotype, y=estimate, fill=Treatment.focus)) +
  geom_errorbar(aes(ymax = estimate + std.error, ymin = estimate - std.error), #, color=Treatment.focus
                width = 0.5, position = position_dodge(width = 0.8)) +
  geom_point(shape=21, size = 2, position = position_dodge(width=0.8)) +
  geom_hline(yintercept = 0, linetype="dotted") 
```

```{r}
gall_selection.df$c.gall_individuals <- with(gall_selection.df, gall_individuals - mean(gall_individuals))

ind_GxE_lms <- lm(gall_survival ~ c.gall_individuals*Genotype*Treatment.focus, data = gall_selection.df, contrast = list(Genotype = "contr.sum"))
summary(ind_GxE_lms)
anova(ind_GxE_lms)
```

```{r}

agg.ind_GxE_lms <- lm(gall_survival ~ (c.gall_individuals + Genotype + Treatment.focus)^2, data = gall_agg, contrast = list(Genotype = "contr.sum"))
summary(agg.ind_GxE_lms)
anova(agg.ind_GxE_lms)
visreg(agg.ind_GxE_lms, xvar="c.gall_individuals", by="Treatment.focus")
visreg(agg.ind_GxE_lms, xvar="c.gall_individuals", by="Genotype")
```

```{r}
gall_selection.df$c.Density_per_100_shoots <- with(gall_selection.df, Density_per_100_shoots - mean(Density_per_100_shoots))

density_GxE_lms <- lm(gall_survival ~ (c.Density_per_100_shoots + Genotype + Treatment.focus)^2, data = gall_selection.df, contrast = list(Genotype = "contr.sum"))
summary(density_GxE_lms)
anova(density_GxE_lms)
```

```{r}
plant_agg <- gall_selection.df %>%
  group_by(Genotype, Treatment.focus, Plant_Position) %>%
  summarise_at(vars(Density_per_100_shoots, gall_survival), funs(mean)) %>%
  ungroup() %>%
  mutate(c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

agg.density_GxE_lms <- lm(gall_survival ~ c.Density_per_100_shoots*Genotype*Treatment.focus, data = plant_agg, contrast = list(Genotype = "contr.sum"))
summary(agg.density_GxE_lms)
anova(agg.density_GxE_lms)
```

```{r}
multivar_GxE_lms <- lm(gall_survival ~ Genotype*Treatment.focus*c.Gall_Height_mm, data = gall_selection.df, contrast = list(Genotype = "contr.sum", Treatment.focus="contr.sum"))
summary(multivar_GxE_lms)
car::vif(multivar_GxE_lms)
anova(multivar_GxE_lms)
visreg(multivar_GxE_lms, xvar="c.Gall_Height_mm", by="Treatment.focus")
visreg(multivar_GxE_lms, xvar="c.gall_individuals", by="Treatment.focus")
visreg(multivar_GxE_lms, xvar="c.Density_per_100_shoots", by="Treatment.focus")
visreg2d(multivar_GxE_lms, xvar="c.Density_per_100_shoots", yvar="c.gall_individuals")
```

```{r Piecewise SEM}
library(lme4)
library(lmerTest)
library(piecewiseSEM)

size_geno <- lm(Gall_Height_mm ~ Genotype, gall_selection.df, contrast = list(Genotype="contr.sum")) 
summary(size_geno)
sem.model.fits(size_geno)
anova(size_geno)

ind_geno <- lm(gall_individuals ~ Genotype, gall_selection.df, contrast = list(Genotype="contr.sum"))
summary(ind_geno)
sem.model.fits(ind_geno)
anova(ind_geno)

density_geno <- lmer(log(Density_per_100_shoots+1) ~ Genotype + (1|Plant_Position/Gall_Number/Gall_ID), gall_selection.df, contrast = list(Genotype="contr.sum"))
summary(density_geno)
sem.model.fits(density_geno)
anova(density_geno)

selection_grads <- lm(gall_survival ~ Treatment.focus*(Gall_Height_mm+gall_individuals+Density_per_100_shoots)^2, gall_selection.df)
sem.model.fits(selection_grads)

model.list <- list(size_geno, ind_geno, density_geno, selection_grads)

sem.fit(model.list, gall_selection.df, corr.errors = c("Gall_Height_mm~~gall_individuals","Gall_Height_mm~~Density_per_100_shoots","gall_individuals~~Density_per_100_shoots"))
```

```{r}
gall_selection.df %>%
  ggplot(., aes(x=c.Gall_Height_mm, y=gall_survival, color=Genotype)) +
  geom_smooth(method="lm") +
  facet_wrap(~Treatment.focus)

gall_agg %>%
  ggplot(., aes(x=c.gall_individuals, y=gall_survival, color=Genotype)) +
  geom_smooth(method="lm") +
  facet_wrap(~Treatment.focus)

plant_agg %>%
  ggplot(., aes(x=c.Density_per_100_shoots, y=gall_survival, color=Genotype)) +
  geom_smooth(method="lm") +
  facet_wrap(~Treatment.focus)
```
